%!TEX root = ../main.tex
\section{Introduction}\label{section:introduction}

The Barab\'asi-Albert (BA) model is a model for generating scale-free networks with a preferential attachment mechanism \citep{Barabasi1999}. It has been extensively studied in many fields, albeit under different names, for its resemblance to real-world networks such as the internet and citation networks, and is the basic model underlying the concept of ``the rich get richer", or more accurately, ``the rich get richer quicker". The fat tail is a central characteristic in the model, which implies that there are a few vertices with very high degree, while most vertices have below average degree. In this model, every new vertex that joins attaches itself to existing vertices with a probability proportional to its degree, so that vertices with a higher degree are more likely to further increase its degree count, leading to a scale free degree distribution. 

There are two key parts to this model: growth and preferential attachment. The random attachment model is a limiting case of the BA model where it retains growth but does not include preferential attachment. Here we show that the resulting degree distribution in this limit is no longer scale-free but geometric, suggesting that growth alone is not enough to produce a power law distribution. 

Recent works \citep{Saramaki2004,Cannings2013,J.P.Saramaki2004} have also investigated if preferential attachment can be reproduced without a global knowledge of the entire network, by performing a random walk on th graph, since such knowledge is impractical for real-world systems. In this project we briefly look at the of such a model and whether it reproduces the power law for preferential attachment. 